Learning features and abstract actions for computing generalized plans

### 2018-09-05 doc/paper2.tex
H's revision at around 23h

\def\year{2019}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai19}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required

@@

\renewcommand\labelitemi{\raisebox{0.4ex}{\tiny$\bullet$}} %% Decrease size of items in enumerate and raise them to center them vertically

@@

\begin{abstract}
  Generalized planning is concerned with the computation of plans that solve multiple instances.
  Recently, it has been shown that generalized plans can be expressed as mappings of feature
  values into actions, and that they can often be computed with fully observable non-deterministic
  (FOND) planners on suitable FOND problems. %after a sequence of transformations.
  The actions in such plans, however, are not the actions in the instances themselves, which are
  not necessarily common to  other instances, but abstract actions that track the effect of the
  actions on the set of features. The formulation assumes that both the features and the abstract
  actions are given.
  In this work, we address this limitation by showing how to automatically learn the features and
  abstract actions. The resulting account of generalized planning combines learning and planning
  in a novel way: a \emph{learner}, based on a min SAT formulation, yields the features and abstract
  actions from state transitions sampled from some problem instances, and a FOND \emph{planner}
  uses this information, suitably transformed, to produce the general plans.
  Correctness guarantees are given and experimental results on several domains are reported.
\end{abstract}

@@

Generalized planning studies the computation of plans for  solving  multiple  instances at once
\cite{srivastava08learning,bonet09automatic,srivastava:generalized,hu:generalized,BelleL16,anders:generalized}.
For example, the plan  that iteratively  picks  a  clear block above $x$ 
and places it  on the table, achieves the goal $clear(x)$
in \emph{any} instance of the Blocksworld where the gripper is initially empty.
Once this  general plan or  policy is derived it can be  applied to solve an infinite collection
of instances that involve different initial states, different objects, and hence  different (ground)  actions.

@@

Bonet and Geffner's  formulation of generalized planning assumes that the
features and abstract actions are given. In this work we address this limitation
by  \emph{showing  how the features and abstract actions  can be learned from
the primitive predicates used to define the instances and a given sample of state transitions.}
For example,  the general policy for achieving $clear(x)$ of \citeay{bonet:ijcai2018}
is obtained using a FOND planner on an abstracton that consists of the boolean
feature $H$ that tracks whether the gripper holds a block and a numerical feature
$n(x)$ that counts the number of blocks above $x$, and two abstract actions:
one with preconditions $\neg H$ and $n(x) > 0$, and effects $H$ and $n(x)\mminus$
(decrement $n(x)$), and the other with precondition $H$ and effect $\neg H$.
In this work, we show how to obtain such policies without having to provide the
features and abstract actions by hand.  The abstraction is learned instead from
the domain predicates and state transitions sampled from some instances.

@@

The paper is organized as follows. We provide the background first and show
how features and abstract actions can be learned by enforcing soundness and
completeness over a set of samples using a pool of features obtained the
domain predicates. The full computational model is then summarized, followed
by experimental results and a discussion.

@@

A \emph{feature} $f$ for a class $\Q$ of problems represents a function $\phi_f$
that takes an instance $P$ from $\Q$ and a state $s$ reachable in $P$, and results
in a value denoted as $\phi_f(s)$. 
A feature is \emph{boolean} if it results in a boolean value, and \emph{numeric}
if it results in a numerical value, that we assume to be a non-negative integer.
The features $H$ and $n(x)$ represent two features in the Blocksworld:
the first is a boolean feature that tracks whether the gripper is empty, while
the second is numerical and tracks the number of blocks above $x$.

@@

An \emph{abstract action}  for a generalized problem $\Q$ and a set of features $F$
is a pair $\bar{a}=\abst{Pre}{\Eff}$ where $Pre$ and $\Eff$ are the  preconditions
and effects expressed in terms of a  set $V_F$ of boolean and numerical \emph{state variables}
associated with the features in $F$. Preconditions and effects over boolean state variables $p$ are literals of the form $p$
and $\neg p$, that abbreviate the  atoms $p=true$ and $p=false$, while  over  numerical state variables $n$,
preconditions are of  the form $n=0$ and $n > 0$, and effects are  of the form $n\mminus$ (decrements) and
$n\pplus$ (increments).

Features $f$ refer to \emph{state functions} $\phi_f(s)$ in the instances
of the generalized problem $\Q$, but to \emph{state variables} in the
abstraction. Likewise, an \emph{abstract state} is  a truth valuation over
the state variables $p$ associated with the boolean features and the atoms
$n=0$ over the state variables $n$ associated with the numerical features.
The abstract state $\bar{s}$ that \emph{corresponds} to a concrete state $s$
is the truth valuation that makes $p$ true iff $\phi_p(s)=true$, and $n=0$
true iff $\phi_n(s) = 0$, and an abstract action $\bar{a}$ is applicable in
a state $s$ if its preconditions are satisfied in $\bar{s}$.

@@

\begin{example}
  For the features $F=\{H,n(x)\}$, the abstract action $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$ 
  is sound in the generalized problem $\Q_{clear}$ which consists of all
  the Blocksworld instances, with stack and unstack actions, whose goal
  is $clear(x)$.
  Indeed, if $s$ is a reachable state in an instance $P$ and $\bar{a}$ is
  applicable in $\bar{s}$, the gripper must be empty in $s$ and some
  block must be above $x$. Thus, if $A$ is the clear block above $x$ and
  $on(A,B)$ is true in $s$, the action $a=unstack(A,B)$ is applicable in
  $s$ and it has the same qualitative effects of $\bar{a}$ over $F$: they
  both make $H$ true and both decrease the value of $n(x)$.
  Likewise, the abstract action $\bar{a'}=\abst{H}{\neg H}$ is also sound. \qed
\end{example}

For a set of abstract actions to be complete in this example, it should 
contain abstracts actions to represent all the concrete actions in the
problem such as stacking a block above $x$, picking up a block that is
not above $x$, and so on.
It may well be possible that a particular feature set $F$ cannot support
a set of abstract actions $A_F$ that is both sound and complete relative
to $\Q$. Soundness, however, is crucial for deriving general plans as it
ensures that in any state of any instance of $\Q$, an applicable abstract
action  can be ``instantiated'' by one or more concrete actions that have
the same qualitative effects on the features.

@@

\begin{enumerate}[1.]
  \item The state variables $V_F$ and the abstract actions $A_F$ are
    extended with initial and goal formulas $I_F$ and $G_F$ over $V_F$
    to yield a \emph{numerical and non-deterministic planning problem}
    $Q_F=\tup{V_F,I_F,G_F,A_F}$. For soundness, $I_F$ must be such that
    the initial states of instances $P$ in $\Q$ all satisfy $I_F$, while
    all states that satisfy $G_F$ must be goal states of $P$.\footnote{A
    state $s$ is said to satisfy $I_F$ or $G_F$ if  $\bar{s}$ does.}
%
  \item The problem $Q_F=\tup{V_F,I_F,G_F,A_F}$ is converted into a
    \emph{boolean FOND} problem $Q'_F=\tup{V'_F,I'_F,G'_F,A'_F}$ by
    replacing the numerical variables $n\in N$ by the symbols $n=0$,
    the first-order literals $n=0$ by propositional literals  $n=0$, 
    the effects $n\pplus$ by deterministic effects $n > 0$, and the
    effects $n\mminus$ by non-deterministic effects $n>0\,|\,n=0$.
%
  \item The solutions computed by a FOND planner on $Q'_F$ are the
    strong cyclic solutions of $Q'_F$ \cite{strong-cyclic}.
    Such solutions however do not necessarily solve $Q_F$ because the
    non-determistic effects $n>0\,|\,n=0$ in $Q'_F$ are \emph{conditionally fair}
    instead of being just \emph{fair}: infinite decrements of $n$ ensure
    $n=0$ eventually \emph{only} when number of increments of $n$ is finite.
    The problem of obtanining solutions of $Q'_F$ that only assume conditional
    fairness, so-called terminating solutions in \cite{srivastava:aaai2011},
    is mapped into the problem of solving an amended FOND $Q^+_F$ that
    only assumes (standard) fairness \cite{bonet:ijcai2017}.\footnote{The
    translation to $Q^+_F$ is performed with software kindly supplied by the authors of \cite{bonet:ijcai2017}.}
    %available by the authors that fixes bugs in the description found
    %in the paper at https://github.com/bonetblai/qnp2fond.}
\end{enumerate}

@@

Bonet and Geffner's formulation and computational model rest on 
features $F$ and  sound abstract actions $A_F$ provided by hand.
The contribution of this work is to learn them both from samples:

\begin{definition}
 For a generalized problem $\Q$, a \emph{sample set}  $\S$ is a non-empty set of tuples  $(s,s',P)$
 such that 1)~$P$ is an instance of $\Q$, 2)~$s$ is a reachable state in $P$, 3)~$s'$
 is a possible succesor of $s$ in $P$; i.e., $s'=f(a,s)$ for some action $a$ applicable in $s$,
 4)~$\S$ is closed in the sense  that if $(s,s',P)$ is in $\S$, then $(s,s'',P)$ is in $\S$  if $s''$ is another possible succesor of $s$ in $P$.
\end{definition}

By assuming that sampled states $s$ are tagged with the instance $P$, 
we abbreviate the tuples $(s,s',P)$ as $(s,s')$, and refer to the
states $s'$ in  the pairs $(s,s') \in \S$ as the successors of
$s$ in $\S$. Notice that the closure condition ensures that
states $s$ appearing \emph{first} in pairs $(s,s')$ must be fully expanded in $\S$;
namely, all possible transitions $(s,s'')$ must be in the sample as well. We call them
the \emph{expanded} states in $\S$.

@@

For a sufficiently large sample set, these approximate notions of soundness and completeness
converge to the exact notions.

@@

For computing the features and abstract actions we define a formula
$T(\S,\F)$ that is \emph{satisfiable} iff there is a set of
features $F \subseteq \F$ and a set of abstract actions $A_F$
over $F$ that is sound and complete relative to the sample set
$\S$.
\textcolor{red}{\bf **** Need to say what $\F$ is ****}

\Omit{The  sets  $F$ and $A_F$ can then be read off from any satisfying
  assignment. Later we will add weights to so me of the clauses to establish
  a preference over these assignments (e.g., for  minimizing $|F|$).
}

@@

For each transition  $(s, s') \in \S$ and each feature $f \in \F$,
$\Delta_f(s, s') \in \{+, -, \uparrow, \downarrow, \bot\}$ will denote the
\emph{qualitative change of value} of feature $f$ along the transition $(s, s')$,
which can go from false to true ($+$), from true to false ($-$) or remain
unchanged ($\bot$), for boolean features, and can increase ($\uparrow$),
decrease ($\downarrow$), or remain unchanged ($\bot$), for numerical features.
The \emph{propositional variables} in $T(\S,\F)$ are then:
\begin{itemize}
  \item $selected(f)$ for each $f \in \F$, true iff $f$ selected (in $F$).
  \item $D_1(s,t)$ for states $s$ and $t$ expanded in $\S$,  true iff
    selected features distinguish $s$ from $t$; i.e.\ if $s$ and $t$ disagree
    on the truth value of a selected boolean feature $p$ or atom $n=0$ for a
    selected numerical feature $n$.
    %Since $D_1$ is symmetric, only symbols for $s < t$ are created for some
    %arbitrary, fixed order over the states in $\S_1$.
  \item $D_2(s, s', t, t')$ for each $(s, s')$ and $(t, t')$ in $\S$,
    true iff selected features $f$ distinguish the two transitions; i.e., if
    $\Delta_f(s, s')\not=\Delta_f(t,t')$. 
    %As for $D_1$, the number of $D_2$ variables is cut in half due to symmetries.
\end{itemize}

@@

\noindent where $f$ ranges over features in $\F$ with  different qualitative
values in $s$ and $t$; i.e.\ boolean features $p$ with different truth value
in $s$ and $t$, and numerical features $n$ with different truth values for
atom $n=0$ in $s$ and $t$.

@@

\noindent where $f$ ranges over features $f \in \F$ that have the same
qualitative values in $s$ and $t$ but which change differently in the two
transitions; i.e., $\Delta_f(s, s') \neq \Delta_f(t, t')$.

The third class of formulas relate $D_1$ and $D_2$ by enforcing \emph{soundness}
and \emph{completeness} over  the samples in $\S$.
Yet, since the abstract actions  are not given but are derived, the qualitative changes in each  transition $(s,s')$ in $\S$
are  taken as templates  of abstract actions. Thus, if $s$ and $t$ are not distinguished by the selected features, 
for each transition $(s, s') \in \S$, there must be a transition $(t, t') \in \S$
such that that the two transitions are not distinguished by the selected features either.
This is expressed as

@@

The theory $T(\S,\F)$ given by the formulas  (\ref{eq:d1})--(\ref{eq:goal})
has a number of atoms that is linear in the size of the feature pool $\F$,
and quadratic in the number of expanded  states $s$  and transitions $(s,s')$ in $\S$.
The resulting number of clauses is \textcolor{red}{\bf XXXX}, which can be
roughly cut in half by  symmetry considerations.

@@

\begin{theorem}
  If $\sigma$ is a satisfying assigment of $T(\S,\F)$, the set of abstract
  actions $A_{\sigma}$ is sound and complete relative to $\S$.
\end{theorem}

@@

\begin{theorem}
For a satisfying assigment $\sigma$  of $T(\S,\F)$, 
$s$ is expanded goal   state in $\S$  iff  $s$ satisfies  $G_F=G_{\sigma}$.
\end{theorem}
}
Different satisfying assignments $\sigma$ give rise to different features $F_{\sigma}$
and abstract actions $A_{\sigma}$. More meaningful features are found by looking for
satisfying assignments $\sigma$ that minimizes a cost measure such as $|F_{\sigma}|$
or more generally $\sum_{f \in F_{\sigma}} cost(f)$. We achieve this constrained
optimization through the use Min SAT or Max-Weighted SAT solvers, where it suffices
to give infinite weights to the clauses resulting from $T(\S,\F)$ and weights
given by $cost(f)$ to the ``soft'' unit clauses $\neg selected(f)$.

@@

The computation of the feature pool $\F$ used in $T(\S,\F)$
relies on the predicates used to encode the instances in $\Q$.
A simple grammar allows  us   to construct new predicates from them, 
and the boolean and numerical features in the pool
are defined from the resulting \emph{unary predicates}. 
The key issue is how to define new predicates from old  ones. For this 
we take a solution off-the-shelf: \emph{description logics (DL)},  also called
\emph{concept languages},  are grammars for doing exactly this.
Concepts $C$  refer  to predicates of arity 1 and roles $R$  to predicates of arity 2.
Concept languages have been used before for  generating general policies using purely learning methods  \cite{martin:concept,fern:bias}.
\citeay{dl-handbook} give syntax and semantics of description logics.

@@

The DL grammar is given by the rules: \textcolor{red}{\bf **** More details?}
\begin{itemize}
  \item $C \ \leftarrow\ \ C_p, C_u, C_x$: primitive, universal, and nominal c's,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
  \item $C \ \leftarrow\  \neg C, C \sqcap C'$: negation and conjunction, 
  \item $C \ \leftarrow\  \exists R.C , \forall R.C$: concepts that use of  roles: first denotes $x$ s.t. for some  $y$,  $R(x,y) \land C(y)$.
  Second,  $x$ s.t. for all $y$, $R(x,y) \rightarrow C(y)$,
  \item $C \ \leftarrow\  R=R'$: stands for $x$ s.t. set of $y$'s  s.t. $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$,
    \item $R \leftarrow R_p, R^{-1}, R^*$: primitive, inverse, transitive closure.
%     \item $R \leftarrow R \circ R'$, role composition: $(x,y)$ in $R \circ R'$ if $(x,z)$ in $R$ and $(z,y)$ in $R'$.
%     \item $R \leftarrow R:C$, denotes pairs $(x,y)$ in $R$ s.t. $C(y)$.
 \end{itemize}

The  compositional  semantics   of these rules allows us to determine the denotations (extensions)   $C^s$ and $R^s$  of
the resulting unary and binary predicates from the denotation of the primitive predicates in any  state $s$.

** Check what is used here and not; see what needs to be explained; keep concise, improve lay out.

@@

The pool $\F=\F^k$  is defined from $\G^k$ and contains
the  features $f$ that represent the functions $\phi_f(s)$ as follows:
\begin{itemize}
\item Boolean $b_p$ for each primitive predicate $p$ of arity $0$,
\item Boolean $b_C$ for each $C$ that are true in $s$ iff $|C^s| > 0$,
\item Numerical $n_C$ for each $C$ that stand for cardinality $|C^s|$,
\item Numerical $\textit{dist}(C_1,R:C,C_2)$, that represent that smallest
  $n$ such that for $x_1, \ldots, x_n$, $C_1^s(x_1)$, $C_2^s(x_{n})$, and
  $(R:C)(x_i,x_{i+1})$ for $i=1, \ldots, n$. 
\end{itemize}

The role $R:C$ is such that its denotation contains the pairs $(x,y)$ in $R$
such that $y$ is in $C$.
Some pruning is done in this space of features as well. E.g., numerical
features $n_C$ are pruned from the pool when there are no states $s$ and $s'$
in the sample set for which $n_c(s)=0$ and $n_c(s') > 1$, while boolean
features $n_b$ are pruned if uniformly true or false for all $s$, or if $n_C$
not pruned. 
The measure $cost(f)$ is set to the complexity of $C$ for $n_C$ and $b_C$,
to $0$ for $b_p$ and to the complexities of $C_1$, $R$, $C$, and $C_2$ added
up for $\textit{dist}(C_1,R:C,C_2)$.
Only features with cost bounded by $k$ are allowed in $\F$.

@@

The policy ...
\alert{Display policy here}

>>>>

\def\year{2019}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai19}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\usepackage{booktabs}  %Required

\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required

@@

@@

\begin{abstract}
  Generalized planning is concerned with the computation of plans that solve multiple instances.
  Recently, it has been shown that generalized plans can be expressed as mappings of feature
  values into actions, and that they can often be computed with fully observable non-deterministic
  (FOND) planners. %  on suitable FOND problems. %after a sequence of transformations.
  The actions in such plans, however, are not the actions in the instances themselves, which are
  not necessarily common to  other instances, but abstract actions that track the effect of 
  actions on the set of features. The formulation assumes that both the features and the abstract
  actions are given.
  In this work, we address this limitation by showing how to learn them automatically. 
  %  the features and  abstract actions.
  The resulting account of generalized planning combines learning and planning
  in a novel way: a \emph{learner}, based on a min SAT formulation, yields the features and abstract
  actions from state transitions sampled from problem instances, and a FOND \emph{planner}
  uses this information, suitably transformed, to produce the general plans.
  Correctness guarantees are given and experimental results on several domains are reported.
\end{abstract}

@@

Generalized planning studies the computation of plans  that   solve   multiple  instances
\cite{srivastava08learning,bonet09automatic,srivastava:generalized,hu:generalized,BelleL16,anders:generalized}.
For example, the plan  that iteratively  picks  a  clear block above $x$ 
and places it  on the table, achieves the goal $clear(x)$
in \emph{any} instance of the Blocksworld where the gripper is initially empty.
Once this  general plan or  policy is derived it can be  applied to solve an infinite collection
of instances that involve different initial states, different objects, and hence  different (ground)  actions.

@@

Bonet and Geffner's  formulation of generalized planning assumes that the
features and abstract actions are given. In this work we address this limitation
by  \emph{showing  how the features and abstract actions  can be learned from
the primitive predicates used to define the instances and a given sample of state transitions.}
For example,  the general policy for achieving $clear(x)$ % of \citeay{bonet:ijcai2018}
is obtained using a FOND planner on an abstraction that consists of  a  boolean
feature $H$  that tracks whether the gripper holds a block,  a numerical feature
$n(x)$ that counts the number of blocks above $x$, and two abstract actions:
one with preconditions $\neg H$ and $n(x) > 0$, and effects $H$ and $n(x)\mminus$
(decrement $n(x)$), and the other with precondition $H$ and effect $\neg H$.
In this work, we show how to obtain such policies from STRIPS instances only
without having to provide the features and abstract actions by hand.

@@

The paper is organized as follows. We provide the background  first and show then
how features and abstract actions can be learned by enforcing soundness and
completeness over a set of samples  using features derived from the domain predicates.
The full computational model is then summarized, followed by experimental results
and a discussion.

@@

A \emph{feature} $f$ for a class $\Q$ of problems represents a function $\phi_f$
that takes an instance $P$ from $\Q$ and a state $s$ reachable in $P$, and results
in a value denoted as $\phi_f(s)$. 
A feature is \emph{boolean} if it results in a boolean value, and \emph{numeric}
if it results in a numerical value, that we assume to be a non-negative integer.
$H$ and $n(x)$ represent two features in the Blocksworld:
the first is a boolean feature that tracks whether the gripper is empty, while
the second is numerical and tracks the number of blocks above $x$.

@@

An \emph{abstract action}  for a generalized problem $\Q$ and a set of features $F$
is a pair $\bar{a}=\abst{Pre}{\Eff}$ where $Pre$ and $\Eff$ are the  preconditions
and effects expressed in terms of a  set $V_F$ of boolean and numerical \emph{state variables}
associated with the features in $F$. Preconditions and effects over boolean state variables $p$ are literals of the form $p$
and $\neg p$, that abbreviate the  atoms $p=true$ and $p=false$, while  over  numerical state variables $n$,
preconditions are of  the form $n=0$ and $n > 0$, and effects  of the form $n\mminus$ (decrements) and
$n\pplus$ (increments).

Features $f$ refer to \emph{state functions} $\phi_f(s)$ in the instances
of the generalized problem $\Q$, but to \emph{state variables} in the
abstraction. Likewise, an \emph{abstract state} is  a truth valuation over
the state variables $p$ associated with the boolean features and the atoms
$n=0$ over the state variables $n$ associated with the numerical features.
The abstract state $\bar{s}$ that \emph{corresponds} to a concrete state $s$
is the truth valuation that makes $p$ true iff $\phi_p(s)=true$, and $n=0$
true iff $\phi_n(s) = 0$. An abstract action $\bar{a}$ is applicable in
a state $s$ if its preconditions are satisfied in $\bar{s}$.

@@

\begin{example}
  For the features $F=\{H,n(x)\}$, the abstract action $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$ 
  is sound in the generalized problem $\Q_{clear}$ which consists of all
  BW instances with stack and unstack actions with goal $clear(x)$.
  Indeed, if $s$ is a reachable state in an instance $P$ and $\bar{a}$ is
  applicable in $\bar{s}$, the gripper must be empty in $s$ and some
  block must be above $x$. Thus, if $A$ is the clear block above $x$ and
  $on(A,B)$ is true in $s$, the action $a=unstack(A,B)$ is applicable in
  $s$ and has the same qualitative effects of $\bar{a}$ over $F$: they
  both make $H$ true and they both decrease the value of $n(x)$.
  Likewise, the abstract action $\bar{a'}=\abst{H}{\neg H}$ is also sound. \qed
\end{example}

For a set of abstract actions to be complete in this example, it must
contain actions to represent all concrete actions in the problem
such as stacking a block above $x$, picking up a block that is
not above $x$, and so on.
It may well be possible that a particular feature set $F$ cannot support
a set of abstract actions $A_F$ that is both sound and complete relative
to $\Q$. Soundness, however, is crucial for deriving general plans as it
ensures that in any state of any instance of $\Q$, an applicable abstract
action  can be ``instantiated'' by one or more concrete actions that have
the same qualitative effects on the features.

@@

\begin{enumerate}[1.]
  \item The state variables $V_F$ and the abstract actions $A_F$ are
    extended with initial and goal formulas $I_F$ and $G_F$ over $V_F$
    to yield a \emph{numerical and non-deterministic planning problem}
    $Q_F=\tup{V_F,I_F,G_F,A_F}$. For soundness, $I_F$ must be such that
    the initial states of instances $P$ in $\Q$ all satisfy $I_F$, while
    all states that satisfy $G_F$ must be goal states of $P$.
    % \footnote{A   state $s$ is said to satisfy $I_F$ or $G_F$ if  $\bar{s}$ does.}
    % no longer needed
%
  \item The problem $Q_F=\tup{V_F,I_F,G_F,A_F}$ is converted into a
    \emph{boolean FOND} problem $Q'_F=\tup{V'_F,I'_F,G'_F,A'_F}$ by
    replacing the numerical variables $n\in N$ by the symbols $n=0$,
    the first-order literals $n=0$ by propositional literals  $n=0$, 
    the effects $n\pplus$ by deterministic effects $n > 0$, and the
    effects $n\mminus$ by non-deterministic effects $n>0\,|\,n=0$.
%
  \item The solutions computed by a FOND planner on $Q'_F$ are the
    strong cyclic solutions of $Q'_F$ \cite{strong-cyclic}.
    Such solutions however do not necessarily solve $Q_F$ because the
    non-deterministic effects $n>0\,|\,n=0$ in $Q'_F$ are not \emph{fair} but
    \emph{conditionally fair}: infinite decrements of $n$ ensure
    $n=0$ eventually \emph{only} when the number of increments of $n$ is finite.
    The problem of obtanining solutions of $Q'_F$ that only assume conditional
    fairness, the so-called terminating solutions in \cite{srivastava:aaai2011},
    is mapped into the problem of solving an amended FOND $Q^+_F$ that
     assumes standard  fairness \cite{bonet:ijcai2017}.\footnote{The
    translation to $Q^+_F$ is performed with software supplied by the authors.}
    %available by the authors that fixes bugs in the description found
    %in the paper at https://github.com/bonetblai/qnp2fond.}
\end{enumerate}

@@

Bonet and Geffner's formulation and computational model rest on 
features $F$ and  sound abstract actions $A_F$ provided by hand.
The contribution of this work is to learn them  from samples:

\begin{definition}
  For a generalized problem $\Q$, a \emph{sample set}  $\S$ is a non-empty set of tuples  $(s,s',P)$
  such that 1)~$P$ is an instance of $\Q$, 2)~$s$ is a reachable state in $P$, 3)~$s'$
  is a possible succesor of $s$ in $P$; i.e., $s'=f(a,s)$ for some action $a$ applicable in $s$,
  4)~$\S$ is closed in the sense  that if $(s,s',P)$ is in $\S$, then $(s,s'',P)$ is in $\S$  if $s''$ is another
  succesor of $s$ in $P$.
\end{definition}

By assuming that sampled states $s$ are tagged with the instance $P$, 
we abbreviate the tuples $(s,s',P)$ as $(s,s')$, and refer to the
states $s'$ in  the pairs $(s,s') \in \S$ as the successors of
$s$ in $\S$. The  closure condition 4)~requires that
states $s$ appearing \emph{first} in pairs $(s,s')$ must be fully expanded in $\S$;
namely, all possible transitions $(s,s'')$ must be in the sample as well. We call them
the \emph{expanded} states in $\S$.

@@

For a sufficiently large sample set, these approximate notions converge to their exact versions.

@@

For computing the features and abstract actions we define a formula
$T(\S,\F)$ where ${\cal S}$ is a sample set for $\Q$ and $\cal F$
is a feature pool defined from the primitive domain predicates
as described below. The formula is  \emph{satisfiable} iff there is a set of
features $F \subseteq \F$ and a set of abstract actions $A_F$
over $F$ that is sound and complete relative to the sample set
$\S$.

@@

For each transition  $(s, s') \in \S$ and each feature $f \in \F$,
$\Delta_f(s, s') \in \{+, -, \uparrow, \downarrow, \bot\}$ will denote the
\emph{qualitative change of value} of feature $f$ along the transition $(s, s')$,
which can go from false to true ($+$), from true to false ($-$) or remain
unchanged ($\bot$), for boolean features, and can increase ($\uparrow$),
decrease ($\downarrow$), or remain unchanged ($\bot$), for numerical features.
The \emph{propositional variables} in $T(\S,\F)$ are then:
\begin{enumerate}[{\small$\bullet$}]
  \item $selected(f)$ for each $f \in \F$, true iff $f$ selected (in $F$).
  \item $D_1(s,t)$ for states $s$ and $t$ expanded in $\S$,  true iff
    selected features distinguish $s$ from $t$; i.e.\ if $s$ and $t$ disagree
    on the truth value of a selected boolean feature $p$ or atom $n=0$ for a
    selected numerical feature $n$.
    %Since $D_1$ is symmetric, only symbols for $s < t$ are created for some
    %arbitrary, fixed order over the states in $\S_1$.
  \item $D_2(s, s', t, t')$ for each $(s, s')$ and $(t, t')$ in $\S$,
    true iff selected features $f$ distinguish the two transitions; i.e., if
    $\Delta_f(s, s')\not=\Delta_f(t,t')$. 
    %As for $D_1$, the number of $D_2$ variables is cut in half due to symmetries.
\end{enumerate}

@@

\noindent where $f$ ranges over the features in $\F$ with  different qualitative
values in $s$ and $t$; i.e.\ boolean features $p$ with different truth value
in $s$ and $t$, and numerical features $n$ with different truth values for
 $n=0$ in $s$ and $t$.

@@

\noindent where $f$ ranges over the features $f \in \F$ that have the same
qualitative values in $s$ and $t$ but which change differently in the two
transitions; i.e., $\Delta_f(s, s') \neq \Delta_f(t, t')$.

The third class of formulas relate $D_1$ and $D_2$ by enforcing \emph{soundness}
and \emph{completeness} over  the samples.  Since the abstract actions  are not given,
the qualitative changes in each  transition $(s,s')$ in $\S$
are  taken as templates  of abstract actions. Thus, if $s$ and $t$ are not distinguished by the selected features, 
for each transition $(s, s') \in \S$, there must be a transition $(t, t') \in \S$
such that that the two transitions are not distinguished by the selected features either.
This is expressed as

@@

The theory $T(\S,\F)$ given by the formulas  (\ref{eq:d1})--(\ref{eq:goal})
has a number of atoms that is linear in the size $|{\cal F}|$  of the feature pool $\F$,
and quadratic in the number $|{\cal S}_1$ of expanded  states $s$  and
in the number $|{\cal S}|_2$ of transitions $(s,s')$ in $\S$.
The resulting number of clauses grows   with
$|{\cal S}_1|^2 \times ({\cal F} + {\cal S}_2)$, and 
can be roughly cut in half by  symmetry considerations.

@@

\begin{theorem}
  If $\sigma$ is a satisfying assigment of $T(\S,\F)$, 
  $A_{\sigma}$ is sound and complete relative to $\S$.
\end{theorem}

@@

\begin{theorem}
For a satisfying assigment $\sigma$  of $T(\S,\F)$, 
$s$ is expanded goal   state in $\S$  iff  $s$ satisfies  $G_F=G_{\sigma}$.
\end{theorem}
}
% 
Different satisfying assignments $\sigma$ give rise to different features $F_{\sigma}$
and abstract actions $A_{\sigma}$. More meaningful features are found by looking for
satisfying assignments $\sigma$ that minimizes a cost measure such as $|F_{\sigma}|$
or more generally $\sum_{f \in F_{\sigma}} cost(f)$. We achieve this
optimization through the use of Min SAT or Max-Weighted SAT solvers, where it suffices
to give infinite weights to the clauses resulting from $T(\S,\F)$ and weights
given by $cost(f)$ to the ``soft'' unit clauses $\neg selected(f)$.

@@

The computation of the feature pool $\F$ used in $T(\S,\F)$
relies on the predicates used to encode the instances in $\Q$.
A simple grammar allows  us   to  construct new predicates 
and to define  the boolean and numerical features in $\F$
in terms of them. The key issue is how to the define
new predicates. For this, we take a solution off-the-shelf: \emph{description logics (DL)},  also called
\emph{concept languages},  are grammars for doing  this.
Concepts $C$  refer  to predicates of arity 1 and roles $R$  to predicates of arity 2.
Concept languages have been used before for  generating general policies using purely learning methods  \cite{martin:concept,fern:bias}.
\citeay{dl-handbook} give syntax and semantics of description logics.

@@

We use a standard  DL grammar  given by the rules:
\begin{enumerate}[{\small$\bullet$}]
  \item $C \ \leftarrow\ \ C_p, C_u, C_x$: primitive, univ.,  nominal concepts,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
  \item $C \ \leftarrow\  \neg C, C \sqcap C'$: negation and conjunction, 
  \item $C \ \leftarrow\  \exists R.C , \forall R.C$: concepts that use of  roles: first denotes $x$ s.t. for some  $y$,  $R(x,y) \land C(y)$.
  Second,  $x$ s.t. for all $y$, $R(x,y) \rightarrow C(y)$,
  \item $C \ \leftarrow\  R=R'$: stands for $x$ s.t. set of $y$'s  s.t. $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$,
    \item $R \leftarrow R_p, R^{-1}, R^*$: primitive, inverse, transitive closure.
%     \item $R \leftarrow R \circ R'$, role composition: $(x,y)$ in $R \circ R'$ if $(x,z)$ in $R$ and $(z,y)$ in $R'$.
%     \item $R \leftarrow R:C$, denotes pairs $(x,y)$ in $R$ s.t. $C(y)$.
 \end{enumerate}

The  compositional  semantics   of these rules allows us to determine the denotations (extensions)   $C^s$ and $R^s$  of
the resulting unary and binary predicates $C^s$ and $R^s$ from the denotation of the primitive predicates
in any  state $s$. We will see examples of derived concepts below such as ${\exists on^* . C_x}$,
whose denotation in a state $s$ of the Blockworld is the set of blocks that are above $x$ in $s$.
The numerical features $n_C$ will be associated with the cardinality of the extension $C^s$.

@@

The pool $\F=\F^k$  is defined from $\G^k$ and contains
the  features $f$ that represent the functions $\phi_f(s)$ as follows:
\begin{enumerate}[{\small$\bullet$}]
\item Boolean $b_p$ for each primitive predicate $p$ of arity $0$,
\item Boolean $b_C$ for each $C$ that are true in $s$ iff $|C^s| > 0$,
\item Numerical $n_C$ for each $C$ that stand for cardinality $|C^s|$,
\item Numerical $\textit{dist}(C_1,R:C,C_2)$, that represent that smallest
  $n$ such that for $x_1, \ldots, x_n$, $C_1^s(x_1)$, $C_2^s(x_{n})$, and
  $(R:C)(x_i,x_{i+1})$ for $i=1, \ldots, n$. 
\end{enumerate}

The role $R:C$ is such that its denotation contains the pairs $(x,y)$ in $R$
such that $y$ is in $C$.
Some pruning is done in this space of features as well. E.g., numerical
features $n_C$ are pruned from the pool when there are no states $s$ and $s'$
in the sample set for which $n_c(s)=0$ and $n_c(s') > 1$, while boolean
features $n_b$ are pruned if uniformly true or false for all $s$, or if $n_C$
is not pruned. 
The measure $cost(f)$ is set to the complexity of $C$ for $n_C$ and $b_C$,
to $0$ for $b_p$ and to the complexities of $C_1$, $R$, $C$, and $C_2$ added
up for $\textit{dist}(C_1,R:C,C_2)$.
Only features with cost bounded by $k$ are allowed in $\F$.

@@

The policy ...
\alert{Display policy here}
}
